---
title: "Project 2: Stats302Project2 Tutorial"
author: "Adam Aley" 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Stats302Project2 Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here is \texttt{Stats302Project2}, a package containing four main functions and tutorials for my Stat 302 Final Project. These four functions include:

- A t-test function
- A linear regression function
- A k-nearest neighbors cross-validation function
- A random forest cross-validation function

## Installation

To download the Stats302Project2 package, use the code below.

Install \texttt{Stats302Project2} using:

```{r, eval = FALSE}
devtools::install_github("Adam-Aley/Stats302Project2")
```

Here, we load our example datasets as a \texttt{gapminder} object:
```{r, message = FALSE}
library(dplyr)
library(tidyr)
library(randomForest)
library(Stats302Project2)
```

# Demonstrating t-test

All the following tests use the lifeExp data from my_gapminder

First, we will demonstrate a test of the hypothesis:
$H_0 : \mu = 60$
$H_A : \mu = 60$
```{r}
# Runs t-test
my_test(my_gapminder$lifeExp, "two-sided", 60)
```

